---
title: "Multiple Factor Analysis"
author: "Jason Sellers"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The mfa package is an R package that implements Multiple Factor Analysis (MFA). For a complete description of MFA, see the paper "Multiple Factor Analysis" (Abdi, Williams, Valentin, 2013).

The working sample consists of a wine tasting experiment, which includes data on 12 wines from three regions (New Zealand, France, and Canada). Ten experts were asked to evaluate these wines on a 9-point rating scale. This package computes the MFA for the wine tasting experiment and provides additional methods for performing various related functions and plotting the results.

## The mfa function

The package includes a main function, mfa(), with the following usage:

```{r}
library(mfa)
df <- read.csv("../data/wines.csv", stringsAsFactors = FALSE)
sets=list(2:7,8:13,14:19)
mfa1 <- mfa(df, sets, ncomps = NULL, center = TRUE, scale = TRUE)
```

where the arguments are:

* **data**: data set (matrix or data frame).
* **sets**: list of vectors indicating the sets of variables (i.e. the blocks).
* **ncomps**: integer indicating how many number of components (i.e. factors) are to be extracted.
* **center**: either a logical value or a numeric vector of length equal to the number of active
variables in the analysis
* **scale**: either a logical value or a numeric vector of length equal to the number of active
variables in the analysis

The mfa function returns an object with the following elements:

* **eigenvalues**: a vector containing the eigenvalues
* **factor_scores**: a matrix of common factor scores
* **partial_factor_scores**: a matrix of partial factor scores
* **loadings**: a matrix of factor loadings

## Plotting Functions

There are three plotting auxiliary functions that can be used with mfa objects:

* **plot.comp**: compromise of the tables
* **plot.pfs**: partial factor scores
* **plot.loadings**: variable loadings

To use each function, simply pass the mfa object to the function, as in the following examples.

```{r}
plot_comp(mfa1)
```

```{r}
plot_pfs(mfa1)
```

```{r}
graphics.off()
plot_loadings(mfa1)
```

## Auxilary Functions

There are five complementary methods and functions which can be used with mfa objects:

* **eigentbl**: table summary of eigenvalues (singular values, eigenvalues, cumulative, % inertia, cumulative % inertia)
* **ob2dim**: contribution of an observation to a dimension
* **var2dim**: contribution of a variable to a dimension
* **tbl2dim**: contribution of a table to a dimension
* **RV**: similarity between two tables using the Rv coefficient
* **RVtbl**: similarity between multiple tables using the Rv coefficient

Each function is used in the following examples:

## eigentbl

```{r}
eigentbl(mfa1)
```

## ob2dim

```{r}
ob2dim(mfa1)
```

## var2dim

```{r}
var2dim(mfa1)
```

## tbl2dim

```{r}
tbl2dim(mfa1)
```

## RV

```{r}
RV(df[,2:7],df[8:13])
```

## RVtbl

```{r}
RVtbl(df,sets=list(2:7,8:13,14:19))
```
